Dissecting Neural Style Transfer

### 2019-06-04 paper/paper.tex
Updates on introduction and related work by Jianzhen

\section{Literature Review}

>>>>

Image stylization is the process that mimics the appeal of artistic media [8] 
and converts input images into a stylized version. It is a research started with
semi-automatic painting systems since 1990s [1] and can be used in many newly 
developed applications in machine learning and artificial intelligence [5]. 
This technique leads to a new form of communication [2]. With the use of image 
stylization, the media users can keep stylistic consistency on media platforms 
like Instagram or Twitter without obtaining skills in traditional artistic 
techniques [16]. Image stylization allows users to generate picture in 
traditional artistic styles, which had been expensive and time-consuming with 
traditional techniques [19]. It shifts the way how the media users share 
pictures on internet. At the beginning of the image stylization, which it 
started in 1980s, procedural filters and patch-based algorithms were studied 
and applied [2]. One example for the procedural filters is to refine image 
with strokes placement algorithms, which a grid of rendering brush strokes is 
placed over an image [13, 3]. Later on, patch-based synthesis was studied based 
on texture recognition [4]. Moreover, these two methods have numerous 
difficulties in further developments on the capture of complex nuances of an 
artist’s style, and can only work for certain types of images [2]. However, 
deep neural networks, which are developed as a general-purpose tool working 
different computer vision problems [5].  The paper published by Leon Gatys in 
2015 has proven that a simple neural network works well on the texture 
recognition just like the patch-based synthesis algorithms [14]. 
One of the applications is auto-painter for cartoon image generation and it 
applies deep neural networks on image stylization [11]. These computational 
bio-inspired neural networks mimic the basic structure of brain neurons. 
Like the auto-painter, convolutional neural network is used in this paper 
for developing a tool in image stylization. Neural networks, especially for 
the deep artificial neural networks, has been widely studied and used in the 
pattern recognition and machine learning in recent years [17]. Convolutional 
neural network is one of the deep learning neural networks and have 
substantially achieved the state-of-the-art accuracies of object recognition 
[12]. This is the key in image stylization where the essential features in the 
style image are extracted and applied on the input image.

This paper applies a pre-trained CNN, VGG-19, on image stylization and 
extends its possibilities on image stylization. CNNs have a similar structure 
as ordinary neural networks, which include neurons with learnable weights and 
biases. CNNs are typically feed-forward architecture and widely used in computer
vision tasks in recent years [12]. Its basic unit is single neuron. Each of 
these neurons takes one or multiple inputs, conducts a dot product then 
optimally follows it with a non-linearity [9]. The core mathematical equation 
for CNN is a result derived from two functions by integration and this indicates 
how one’s shape is changed by the other. CNN works in a way similar to broad 
categorization [19]. When a CNN is used in image processing, every pixel of the 
input image will be processed by a feature detector and a feature map is given 
as the result. There will be very less, but critical features kept in the 
feature map compared to the input image. Consequently, some key features might 
be missed, and massive trainings are required for CNN’s accuracy on texture 
capture.

\section{Related Work}

In the field of image stylization, there are abundant related work found.

2.1 Image stylization using stroke-based and example-based methods

Both stroke-based and example-based methods have been studied since last century 
for image stylization. For the stroke-based stylization, one typical method is 
that brush strokes are aligned based on input image for varying color, size and 
orientation [18]. Many methods, which are developed based on brush strokes, 
normally either focus on one specific type of brush or provides a generic 
framework [7]. For an instance, a brush stroke for a traditional Japanese 
sumi-e style painting creation is modelled with the use of a 1D array of 
bristles and this method can only generate a stylized image in grey scale [20]. 
One major disadvantage of the stroke-based method is that it cannot generate 
result automatically and provides tools to users as they can tweak the output 
[7]. Many researches have also pointed out that the complex nuances of an 
artist’s style are too hard to be captured in some degrees [2]. As for the 
example-based methods, texture recognition and transfer are the core parts [18].
One framework for this method is called the image analogies, which is a basic 
reasoning process and consisted of two stages: a design phase with training 
data and application phase with new images [6]. This method is still being 
studied but is more complex than the neural network [2].

2.2 Image stylization using convolutional neural networks

An extension to the example-based method would be the texture synthesis using 
convolutional neural networks developed by Leon Gatys [14]. This is the 
pre-trained CNN used in the paper, which is also known as VGG-19. Gram matrices 
are introduced on the feature maps to transfer the representations into a 
stationary feature space [14]. There are also other works that have used 
convolutional neural networks for image stylization [10, 15]. In the process of 
image stylization, there is a trade-off between content and style matching, 
which is the content loss [10]. Furthermore, there are some limitations of 
using VGG networks due to the zero-padding used for convolutions in VGG network 
[14]. Overall, convolutional neural networks have a better performance than 
previous works on image stylization [21].

References for intro and related work:
[1] Haeberli, P. (1990). Paint by numbers: abstract image representations. SIGGRAPH Comput Graph, 24 (4), pp. 207-214. doi: 10.1145/97880.97902
[2] Hertzmann, A. (2018). Image Stylization: History and Future (Part1). Adobe Research. Viewed at: https://research.adobe.com/news/image-stylization-history-and-future/
[3] Hertzmann, A. (1998). Painterly rendering with curved brush strokes of multiple sizes. Proceedings of the 25th annual conference on Computer graphics and interactive techniques, pp. 453-460. doi: 10.1145/280814.280951 
[4] Hertzmann, A. (2018). Image Stylization: History and Future (Part2). Adobe Research. Viewed at: https://research.adobe.com/news/image-stylization-history-and-future-part-2/
[5] Hertzmann, A. (2018). Image Stylization: History and Future (Part3). Adobe Research. Viewed at: https://research.adobe.com/news/image-stylization-history-and-future-part-3/
[6] Hertzmann, A., Jacobs, C. E., Oliver, N., Curless, B., Salesin, D. H. (2001). Image analogies. Proceedings of the 28th annual conference on Computer graphics and interactive techniques, pp. 327-340. doi: 10.1145/383259.383295 
[7] Hegde, S., Gatzidis, C., Tian, F. (2012). Painterly rendering techniques: a state‐of‐the‐art review of current approaches. Wiley Online Library. doi: 10.1002/cav.1435
[8] Kyprianidis, J.E., Collomosse, J., Wang, T., Isenberg, T. (2013). State of the art: a taxonomy of artistic stylization techniques for images and video. IEEE Trans Vis Comput Graphics, 19 (5), pp. 866-885. doi: 10.1109/TVCG.2012.160
[9] Karpathy, A. (2019). Convolutional Neural Networks (CNNs/ConvNets). Stanford, CS231n Convolutional Neural Networks for Visual Recognition. Viewed at:  http://cs231n.github.io/convolutional-networks/
[10] Krizhevsky, A., Sutskever, I., Hinton, G. E. (2012). ImageNet Classification with Deep Convolutional Neural Networks. NIPS proceedings. Viewed at: http://papers.nips.cc/paper/4824-imagenet-classification-with-deep-convolutional-neural-networ
[11] Liu, Y., Qin, Z., Wan, T., Luo, Z. (2018). Auto-painter: Cartoon image generation from sketch by using conditional Wasserstein generative adversarial networks. Neurocomputing, Volume 311, pp. 78-87. ISSN 0925-2312. doi: 10.1016/j.neucom.2018.05.045. 
[12] Liang, M., Hu, X. (2015). Recurrent Convolutional Neural Network for Object Recognition. IEEE Conference on Computer Vision and Pattern Recognition (CVPR), pp. 3367-3375. Viewed at: http://openaccess.thecvf.com/content_cvpr_2015/html/Liang_Recurrent_Convolutional_Neural_2015_CVPR_paper.html
[13] Litwinowicz, P. (1997). Processing images and video for an impressionist effect. Apple Computer, Inc., CA, DOI: 10.1145/258734.258893
[14] Gatys, L. A., Ecker, A. S., Bethge, M. (2015), Texture Synthesis using Convolutional Neural Networks. Advances in Neural Information Processing Sytems 28 (NIPS 2015), Viewed at: http://papers.nips.cc/paper/5633-texture-synthesis-using-convolutional-neural-networks
[15] Gatys, L. A., Ecker, A. S., Bethge, M. (2016). Image Style Transfer Using Convolutional Neural Networks. IEEE Conference on Computer Vision and Pattern Recognition (CVPR), pp. 2414-2423. Viewed at: https://www.cv-foundation.org/openaccess/content_cvpr_2016/html/Gatys_Image_Style_Transfer_CVPR_2016_paper.html
[16] Rosin, P., Collomosse, J. (2013). Image and Video-Based Artistic Stylisation. Springer. doi: 10.1007/978-1-4471-4519-6 
[17] Schmidhuber, J. (2014). Deep learning in neural networks: An overview. Neural Networks, Volume 61, pp. 85-117. ISSN 0893-6080. doi: 10.1016/j.neunet.2014.09.003.
[18] Semmo, A., Limberger, D., Kyprianidis, J. E., Döllner, J. (2016). Image stylization by interactive oil paint filtering. Computers & Graphics, Volume 55, pp. 157-171, ISSN 0097-8493, doi: 10.1016/j.cag.2015.12.001.
[19] Super Data Science (2018). Convolutional Neural Networks (CNN): Step 1 – Convolution Operation. Super Data Science. Viewed at: https://www.superdatascience.com/blogs/convolutional-neural-networks-cnn-step-1-convolution-operation/ 
[20] Strassmann, S. (1986). Hairy brushes. Proceedings of the 13th Annual Conference on Computer Graphics and Interactive Techniques 1986, 20(4), pp.225–232.
[21] Zhu, F., Yu, Y. (2018). Automatic Image Stylization using Deep Fully Convolutional Networks. The University of Hong Kong. Viewed at: https://arxiv.org/pdf/1811.10872.pdf
